---
title: "Untitled"
author: "Unnat Palan"
date: "August 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(caret)
library(rpart)
library(randomForest)
library(e1071)
library(gbm)
```

```{r}
set.seed(1234)

TrainDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
TestDataURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
TrainingData <- read.csv(url(TrainDataURL), na.strings=c("NA","#DIV/0!",""))
TestingData <- read.csv(url(TestDataURL), na.strings=c("NA","#DIV/0!",""))
dim(TrainingData)
dim(TestingData)
```

Deleting columns with missing values

```{r}
TrainingData <-TrainingData[,colSums(is.na(TrainingData)) == 0]
TestingData <-TestingData[,colSums(is.na(TestingData)) == 0]
```

Deleting unused columns

```{r}
TrainingData <-TrainingData[,-c(1:7)]
TestingData <-TestingData[,-c(1:7)]
dim(TrainingData)
dim(TestingData)
```

## Dataset Partitioning

```{r}
SplitData <- createDataPartition(TrainingData$classe, p=0.7, list=FALSE)
TrainingSet <- TrainingData[SplitData, ]
TestingSet <- TrainingData[-SplitData, ]
dim(TrainingSet)
dim(TestingSet)
```

## Examining the data based on it's frequency

```{r}
plot(TrainingSet$classe, col="blue", main="Frequency of different levels", xlab="classe", ylab="Frequency")
```

## Prediction using Random Forest and Generalized Boosted Models

- Random Forest

```{r}
rfmodel <- randomForest(classe ~., data=TrainingSet, method="class")
rfprediction <- predict(rfmodel, TestingSet, Type="class")
confusionMatrix(rfprediction, TestingSet$classe)
```

- Generalized Boosted Model

```{r}
fitControl <- trainControl(method="repeatedcv", number=5, repeats=1)
gbmodel <- train(classe ~., data=TrainingSet, method="gbm", trControl=fitControl, verbose=FALSE)
gbprediction <- predict(gbmodel, TestingSet)
confusionMatrix(gbprediction, TestingSet$classe)
```

The accuracy of the Random Forest Model is 99.64, which is higher than that of the  Generalized Boosted Model, hence the Random Forest model will be applied to predict the 20 quiz results.

```{r}
predictTest <- predict(rfmodel, TestingData)
predictTest
```

## Submission 

Creating the file for submission

```{r}
pml_write_files = function(x){
n = length(x)
for(i in 1:n){
filename = paste0("problem_id_",i,".txt")
write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
}
}
pml_write_files(predictTest)
```
